---
title: "Histograms of Power Differences within Electrodes"
author: Lindsay Vass
date: 28 May 2015
output: html_document
---

# Calculation of Power

Six-cycle wavelets were used to estimate the power at each frequency and time
point of the cleaned data set. I then took the log(Power) and z-scored these
values within each electrode and frequency. All plotted values are of these
z-scored log(Power) values.

* 31 log-spaced frequencies between 1 Hz and 181 Hz
* 6 cycle wavelet for power estimation

I then extracted the mean power value for each trial in a series of 
contiguous time bins that surrounded the teleportation event. These bins were
defined as follows (t0 = time of teleporter entry):

* -1000 : 0 ms
* Time in Teleporter (0 : 1830 ms for short time; 0 : 2830 ms for long time)
* Exit time : Exit time + 1000 ms


``` {r, echo = FALSE, message = FALSE}
library(dplyr)
library(reshape2)
library(permute)
library(ggplot2)

setwd('/Users/Lindsay/Documents/MATLAB/iEEG/Group Analysis/Power/')
load("All_Subjects_Power_CLEAN_3sBuffer2015-05-27.Rda")
```


# Power Differences by Timepoint

In this first analysis, I collapsed across the spatial and temporal conditions
for each trial to examine whether there were significant power differences
between timepoints within a trial. I used a Wilcoxon signed-rank test to 
evaluate the significance of the following four comparisons:

* Pre > Tele
* Pre < Tele
* Tele > Post
* Tele < Post

For each comparison, I counted the number of electrodes that attained P < 0.05
according to the Wilcoxon signed-rank test. I then used a one-sided binomial
test to determine whether the total number of significant electrodes was greater
than would be expected by chance (`r nlevels(allPowerData$ElectrodeID)` electrodes
x 5% = `r nlevels(allPowerData$ElectrodeID) * 0.05` electrodes). Comparisons
that resulted in a significant binomial test (P < 0.05) are signified with an
asterisk (*). 

``` {r, echo = FALSE, message = FALSE, fig.width = 20, fig.height = 10}
# Initialize output data frame and prepare input data frame ---------------

wilcoxonResults <- data.frame(ElectrodeID = NA,
                              FrequencyBand = NA,
                              Contrast = NA,
                              P = NA)
thisRow <- 1

# Prepare input data
allPowerData <- allPowerData %>%
  mutate(TrialID = paste(EDF, TrialNumber, sep = "-")) %>% # Add a variable for TrialID since there can be a trial #1 for EDF1 and a trial #1 for EDF2
  filter(TimePoint == "Pre1" | TimePoint == "Tele" | TimePoint == "Post1") %>% # keep only timepoints of interest
  select(TimePoint, Power:TrialID) 

# Perform analysis for each electrode x frequency band --------------------

for (thisElectrode in 1:nlevels(allPowerData$ElectrodeID)){
  
  # Get all data for this electrode
  electrodeData <- allPowerData %>%
    filter(ElectrodeID == levels(ElectrodeID)[thisElectrode]) %>%
    select(-ElectrodeID) 
  
  for (thisFreqBand in 1:nlevels(allPowerData$FrequencyBand)) {
    
    # Filter data for this frequency band
    frequencyData <- electrodeData %>%
      filter(FrequencyBand == levels(FrequencyBand)[thisFreqBand]) %>%
      select(-FrequencyBand) %>%
      group_by(TrialID, TimePoint) %>%
      summarise(Power = mean(Power))
    
    # Cast to wide format
    frequencyDataWide <- frequencyData %>%
      dcast(TrialID ~ TimePoint, value.var = "Power")
    
    # Wilcoxon signed-rank tests
    preGtTele  <- wilcox.test(frequencyDataWide$Pre1, frequencyDataWide$Tele, alternative = "greater", paired=TRUE)
    preLtTele  <- wilcox.test(frequencyDataWide$Pre1, frequencyDataWide$Tele, alternative = "less", paired=TRUE)
    teleGtPost <- wilcox.test(frequencyDataWide$Tele, frequencyDataWide$Post1, alternative = "greater", paired = TRUE)
    teleLtPost <- wilcox.test(frequencyDataWide$Tele, frequencyDataWide$Post1, alternative = "less", paired = TRUE)
    
    # Add to summary dataframe
    wilcoxonResults[thisRow, ]     <- c(levels(allPowerData$ElectrodeID)[thisElectrode], 
                                      levels(allPowerData$FrequencyBand)[thisFreqBand], 
                                      "Pre > Tele", 
                                      preGtTele$p.value)
    wilcoxonResults[thisRow + 1, ] <- c(levels(allPowerData$ElectrodeID)[thisElectrode], 
                                      levels(allPowerData$FrequencyBand)[thisFreqBand], 
                                      "Pre < Tele", 
                                      preLtTele$p.value)
    wilcoxonResults[thisRow + 2, ] <- c(levels(allPowerData$ElectrodeID)[thisElectrode], 
                                      levels(allPowerData$FrequencyBand)[thisFreqBand], 
                                      "Tele > Post", 
                                      teleGtPost$p.value)
    wilcoxonResults[thisRow + 3, ] <- c(levels(allPowerData$ElectrodeID)[thisElectrode], 
                                      levels(allPowerData$FrequencyBand)[thisFreqBand], 
                                      "Tele < Post", 
                                      teleLtPost$p.value)
    thisRow <- thisRow + 4
    
  } # end thisFreqBand
  
} # end thisElectrode


# Summarize significant wilcoxon results ----------------------------------

# Reorder frequency band levels
freqBandOrder <- c("Delta", "Theta", "Alpha", "Beta", "Gamma")
wilcoxonResults$FrequencyBand <- factor(wilcoxonResults$FrequencyBand, levels = freqBandOrder)

# Get the counts of the significant results
wilcoxonSigResults <- wilcoxonResults %>%
  filter(P < 0.05) %>%
  group_by(FrequencyBand, Contrast) %>%
  summarise(Count = n())

# Perform binomial test to determine whether counts are higher than expected by chance
binomP <- wilcoxonSigResults %>%
  rowwise() %>%
  do(BinomialTestP = binom.test(.$Count, 
                                nlevels(allPowerData$ElectrodeID), 
                                p = 0.05, 
                                alternative = "greater")$p.value)
wilcoxonSigResults <- cbind(wilcoxonSigResults, binomP) 

# Make a dataframe for the significance markers
sigMarkers <- wilcoxonSigResults %>%
  filter(BinomialTestP < 0.05) %>%
  mutate(Count = Count + 0.2) 

# Plot the bar charts
wilcoxonPlot <- wilcoxonSigResults %>%
  ggplot(aes(x = Contrast, y = Count)) +
  geom_bar(stat = "identity") +
  geom_hline(aes(yintercept = 0.05*nlevels(allPowerData$ElectrodeID), 
                 color = "red")) +
  facet_grid(. ~ FrequencyBand) +
  geom_text(data = sigMarkers, label = "*", size = 18) +
  theme(strip.text.x = element_text(size = 18),
        plot.title = element_text(size = 24, vjust = 2),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 18),
        axis.text.x = element_text(size = 18, angle = 90)) +
  ggtitle("Frequency of Significant Differences in Z-scored log(Power) by Frequency Band")
wilcoxonPlot
```

# Power Differences by Spatial Condition

In this second analysis, I collapsed across the temporal conditions and timepoints
for each trial to examine whether there were significant power differences
between spatial conditions across trials. I had initially planned on a 2-way
Space x Timepoint ANOVA, but I realized that this would require a mixed model
(Repeated = Timepoint, Random = Space) and I'm not comfortable enough with
covariance structures to implement that at this time. I used a Mann-Whitney U
test to evaluate the significance of the following two comparisons:

* NS > FS
* NS < FS

For each comparison, I counted the number of electrodes that attained P < 0.05
according to the Wilcoxon signed-rank test. I then used a one-sided binomial
test to determine whether the total number of significant electrodes was greater
than would be expected by chance (`r nlevels(allPowerData$ElectrodeID)` electrodes
x 5% = `r nlevels(allPowerData$ElectrodeID) * 0.05` electrodes). Comparisons
that resulted in a significant binomial test (P < 0.05) are signified with an
asterisk (*). 

``` {r, echo = FALSE, message = FALSE, fig.width = 20, fig.height = 10}

# Refresh original version of raw data
load("All_Subjects_Power_CLEAN_3sBuffer2015-05-27.Rda")

# Initialize output data frame and prepare input data frame ---------------

wilcoxonResults <- data.frame(ElectrodeID = NA,
                              FrequencyBand = NA,
                              Contrast = NA,
                              P = NA)
# Prepare input data
allPowerData <- allPowerData %>%
  mutate(TrialID = paste(EDF, TrialNumber, sep = "-")) %>% # Add a variable for TrialID since there can be a trial #1 for EDF1 and a trial #1 for EDF2
  filter(TimePoint == "Pre1" | TimePoint == "Tele" | TimePoint == "Post1") %>% # keep only timepoints of interest
  select(TrialSpaceType, TimePoint, Power:TrialID) 

# Perform analysis for each electrode x frequency band --------------------

for (thisElectrode in 1:nlevels(allPowerData$ElectrodeID)){
  
  # Get all data for this electrode
  electrodeData <- allPowerData %>%
    filter(ElectrodeID == levels(ElectrodeID)[thisElectrode]) %>%
    select(-ElectrodeID) 
  
  for (thisFreqBand in 1:nlevels(allPowerData$FrequencyBand)) {
    
    # Filter data for this frequency band
    frequencyData <- electrodeData %>%
      filter(FrequencyBand == levels(FrequencyBand)[thisFreqBand]) %>%
      select(-FrequencyBand) %>%
      group_by(TrialID, TrialSpaceType) %>%
      summarise(Power = mean(Power))
    
    # Cast to wide format
    frequencyDataWide <- frequencyData %>%
      dcast(TrialID ~ TrialSpaceType, value.var = "Power")
    
    # Wilcoxon signed-rank tests
    NSGtFS  <- wilcox.test(frequencyDataWide$NS, frequencyDataWide$FS, alternative = "greater")
    NSLtFS  <- wilcox.test(frequencyDataWide$NS, frequencyDataWide$FS, alternative = "less")
    
    # Add to summary dataframe
    wilcoxonResults[thisRow, ]     <- c(levels(allPowerData$ElectrodeID)[thisElectrode], 
                                      levels(allPowerData$FrequencyBand)[thisFreqBand], 
                                      "NS > FS", 
                                      NSGtFS$p.value)
    wilcoxonResults[thisRow + 1, ] <- c(levels(allPowerData$ElectrodeID)[thisElectrode], 
                                      levels(allPowerData$FrequencyBand)[thisFreqBand], 
                                      "NS < FS", 
                                      NSLtFS$p.value)
    thisRow <- thisRow + 2
    
  } # end thisFreqBand
  
} # end thisElectrode


# Summarize significant wilcoxon results ----------------------------------

# Reorder frequency band levels
freqBandOrder <- c("Delta", "Theta", "Alpha", "Beta", "Gamma")
wilcoxonResults$FrequencyBand <- factor(wilcoxonResults$FrequencyBand, levels = freqBandOrder)

# Get the counts of the significant results
wilcoxonSigResults <- wilcoxonResults %>%
  filter(P < 0.05) %>%
  group_by(FrequencyBand, Contrast) %>%
  summarise(Count = n())

# Perform binomial test to determine whether counts are higher than expected by chance
binomP <- wilcoxonSigResults %>%
  rowwise() %>%
  do(BinomialTestP = binom.test(.$Count, 
                                nlevels(allPowerData$ElectrodeID), 
                                p = 0.05, 
                                alternative = "greater")$p.value)
wilcoxonSigResults <- cbind(wilcoxonSigResults, binomP) 

# Make a dataframe for the significance markers
sigMarkers <- wilcoxonSigResults %>%
  filter(BinomialTestP < 0.05) %>%
  mutate(Count = Count + 0.2) 

# Plot the bar charts
wilcoxonPlot <- wilcoxonSigResults %>%
  ggplot(aes(x = Contrast, y = Count)) +
  geom_bar(stat = "identity") +
  geom_hline(aes(yintercept = 0.05*nlevels(allPowerData$ElectrodeID), 
                 color = "red")) +
  facet_grid(. ~ FrequencyBand) +
  geom_text(data = sigMarkers, label = "*", size = 18) +
  theme(strip.text.x = element_text(size = 18),
        plot.title = element_text(size = 24, vjust = 2),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 18),
        axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18)) +
  ggtitle("Frequency of Significant Differences in Z-scored log(Power) by Frequency Band")
wilcoxonPlot
```

# Power Differences by Temporal Condition

In this third analysis, I collapsed across the spatial conditions and timepoints
for each trial to examine whether there were significant power differences
between temporal conditions across trials. (Same reasoning regarding two-way
ANOVA as above for Spatial Condition.) I used a Mann-Whitney U test to evaluate 
the significance of the following two comparisons:

* NT > FT
* NT < FT

For each comparison, I counted the number of electrodes that attained P < 0.05
according to the Wilcoxon signed-rank test. I then used a one-sided binomial
test to determine whether the total number of significant electrodes was greater
than would be expected by chance (`r nlevels(allPowerData$ElectrodeID)` electrodes
x 5% = `r nlevels(allPowerData$ElectrodeID) * 0.05` electrodes). Comparisons
that resulted in a significant binomial test (P < 0.05) are signified with an
asterisk (*). 

``` {r, echo = FALSE, message = FALSE, fig.width = 20, fig.height = 10}

# Refresh original version of raw data
load("All_Subjects_Power_CLEAN_3sBuffer2015-05-27.Rda")

# Initialize output data frame and prepare input data frame ---------------

wilcoxonResults <- data.frame(ElectrodeID = NA,
                              FrequencyBand = NA,
                              Contrast = NA,
                              P = NA)
thisRow <- 1

# Prepare input data
allPowerData <- allPowerData %>%
  mutate(TrialID = paste(EDF, TrialNumber, sep = "-")) %>% # Add a variable for TrialID since there can be a trial #1 for EDF1 and a trial #1 for EDF2
  filter(TimePoint == "Pre1" | TimePoint == "Tele" | TimePoint == "Post1") %>% # keep only timepoints of interest
  select(TrialTimeType, TimePoint, Power:TrialID) 

# Perform analysis for each electrode x frequency band --------------------

for (thisElectrode in 1:nlevels(allPowerData$ElectrodeID)){
  
  # Get all data for this electrode
  electrodeData <- allPowerData %>%
    filter(ElectrodeID == levels(ElectrodeID)[thisElectrode]) %>%
    select(-ElectrodeID) 
  
  for (thisFreqBand in 1:nlevels(allPowerData$FrequencyBand)) {
    
    # Filter data for this frequency band
    frequencyData <- electrodeData %>%
      filter(FrequencyBand == levels(FrequencyBand)[thisFreqBand]) %>%
      select(-FrequencyBand) %>%
      group_by(TrialID, TrialTimeType) %>%
      summarise(Power = mean(Power))
    
    # Cast to wide format
    frequencyDataWide <- frequencyData %>%
      dcast(TrialID ~ TrialTimeType, value.var = "Power")
    
    # Wilcoxon signed-rank tests
    NTGtFT  <- wilcox.test(frequencyDataWide$NT, frequencyDataWide$FT, alternative = "greater")
    NTLtFT  <- wilcox.test(frequencyDataWide$NT, frequencyDataWide$FT, alternative = "less")
    
    # Add to summary dataframe
    wilcoxonResults[thisRow, ]     <- c(levels(allPowerData$ElectrodeID)[thisElectrode], 
                                      levels(allPowerData$FrequencyBand)[thisFreqBand], 
                                      "NT > FT", 
                                      NTGtFT$p.value)
    wilcoxonResults[thisRow + 1, ] <- c(levels(allPowerData$ElectrodeID)[thisElectrode], 
                                      levels(allPowerData$FrequencyBand)[thisFreqBand], 
                                      "NT < FT", 
                                      NTLtFT$p.value)
    thisRow <- thisRow + 2
    
  } # end thisFreqBand
  
} # end thisElectrode


# Summarize significant wilcoxon results ----------------------------------

# Reorder frequency band levels
freqBandOrder <- c("Delta", "Theta", "Alpha", "Beta", "Gamma")
wilcoxonResults$FrequencyBand <- factor(wilcoxonResults$FrequencyBand, levels = freqBandOrder)

# Get the counts of the significant results
wilcoxonSigResults <- wilcoxonResults %>%
  filter(P < 0.05) %>%
  group_by(FrequencyBand, Contrast) %>%
  summarise(Count = n())

# Perform binomial test to determine whether counts are higher than expected by chance
binomP <- wilcoxonSigResults %>%
  rowwise() %>%
  do(BinomialTestP = binom.test(.$Count, 
                                nlevels(allPowerData$ElectrodeID), 
                                p = 0.05, 
                                alternative = "greater")$p.value)
wilcoxonSigResults <- cbind(wilcoxonSigResults, binomP) 

# Make a dataframe for the significance markers
sigMarkers <- wilcoxonSigResults %>%
  filter(BinomialTestP < 0.05) %>%
  mutate(Count = Count + 0.2) 

# Plot the bar charts
wilcoxonPlot <- wilcoxonSigResults %>%
  ggplot(aes(x = Contrast, y = Count)) +
  geom_bar(stat = "identity") +
  geom_hline(aes(yintercept = 0.05*nlevels(allPowerData$ElectrodeID), 
                 color = "red")) +
  facet_grid(. ~ FrequencyBand) +
  geom_text(data = sigMarkers, label = "*", size = 18) +
  theme(strip.text.x = element_text(size = 18),
        plot.title = element_text(size = 24, vjust = 2),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 18),
        axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18)) +
  ggtitle("Frequency of Significant Differences in Z-scored log(Power) by Frequency Band")
wilcoxonPlot
```

