---
title: "Pepisode during teleporter does not predict latency to enter the correct arm on the next trial"
author: "Lindsay Vass"
date: "22 June 2015"
output: html_document
---

## Calculation of Pepisode Values

#### Building the power distribution
First I built the distribution of power values by taking the cleaned data and 
calculating the power at each frequency at every time point of the entire navigation
period when subjects were searching for stores.

Parameters:

* 31 log-spaced frequencies between 1 Hz and 181 Hz
* 6 cycle Morlet wavelets

#### Epoching the data

Then, I extracted epoched data for each trial. The epochs were defined as follows:

* Start time: 3000 ms before teleporter entry
* End time: 3000 ms after teleporter exit

#### Pepisode calculation

I calculated the power at each frequency at every time point of the epoch, using
mirrored buffers (length = 3 cycles at 1 Hz) to avoid edge artifacts, again
using 6 cycle Morlet wavelets.

This timeseries data was then amplitude-thresholded using the value at 95% of the
distribution calculated earlier. This returns a binary vector which indicates
whether power exceeded the amplitude threshold at each time point. This vector
was then duration-thresholded (n >= 3 cycles) so that only episodes lasting at
least 3 cycles are retained. The result is a binary vector showing times when
significant oscillatory episodes are present.

#### Re-epoching the data

After calculating pepisode at each time point of the trial, the Pre and Post
time bins were trimmed to be the same length as the Teleporter bin (1830 ms
for NT trials and 2830 ms for FT trials).

## Calculation of latency

The value we are attempting to predict is the latency to enter the correct arm
of the maze. This is defined as the time at which the patient crosses the 
boundary between the central plaza and the correct arm of the maze. Trials for
which the patient exited the teleporter facing the correct arm (1/3 of trials)
were excluded from the analysis.

## Scatterplot of Pepisode vs. Latency

The plot below shows the relationship between pepisode on trial N and the 
latency to enter the correct arm of the maze on trial N+1. There does not appear
to be a relationship between these values for any electrode.

```{r, echo = FALSE, message = FALSE, fig.width=16, fig.height=12}
library(dplyr)
library(ggplot2)
load('../Rda/allAnalyzedData.Rda')

for (thisFreqBand in 1:nlevels(validData$FrequencyBand)) {
  p <- validData %>%
  filter(validData$FrequencyBand == levels(validData$FrequencyBand)[thisFreqBand]) %>%
  ggplot(aes(x = PrevTrialPepisode, y = Latency)) +
  geom_point() +
  facet_wrap(~ElectrodeID) +
  ggtitle(paste(levels(validData$FrequencyBand)[thisFreqBand], "Pepisode"))
  print(p)
}


```

## Modeling Pepisode vs. Latency Within-Electrode

Each electrode with at least 10 valid trials (N = `r nlevels(glmOutput$ElectrodeID)`) 
was included in the analysis. Data for each frequency band were analyzed 
separately. The dependent variable (y) was the latency to enter the correct arm 
of the maze. The independent variables (x) included: 

* Pepisode during the teleporter on the previous trial
* Previous trial's time type (NT/FT)
* Previous trial's space type (NS/FS)

The data were fit with a linear model of the quasipoisson family since latency
is bounded by 0. To determine the true P value for the Beta estimates, the data
for each model were randomized `r numPerm` times to build a distribution of Beta
estimates. Because pepisode should show a negative relationship with latency --
that is, greater pepisode should result in lower latency -- the corrected P value
was defined based on the number of observations in the distribution that were
more negative than the observed Beta estimate. The P value threshold was then adjusted
using the Bonferroni correction for the number of frequency bands analyzed (N = 
`r nlevels(validData$FrequencyBand)`; P < `r 0.05/nlevels(validData$FrequencyBand)`).

```{r, echo = FALSE}

glmOutput$FrequencyBand <- factor(glmOutput$FrequencyBand)
correctedOutput <- glmOutput %>%
  filter(term == "PrevTrialPepisode" &
           CorrectedP < 0.05 / nlevels(FrequencyBand)) %>%
  select(ElectrodeID, FrequencyBand, CorrectedP) %>%
  rename(PValue = CorrectedP)
print(correctedOutput)

```

As shown above, across all frequency bands, there were only `r nrow(correctedOutput)`
 observations that exceeded the Bonferroni-corrected threshold. All of these
observations were from the same patient and on the same depth electrode.