---
title: "Pepisode Timing Results"
author: "Lindsay Vass"
date: "9 June 2015"
output: html_document
---

## Calculation of Pepisode Values

First I built the distribution of power values by taking the cleaned data and 
calculating the power at each frequency at every time point of the entire navigation
period when subjects were searching for stores.

Parameters:

* 31 log-spaced frequencies between 1 Hz and 181 Hz
* 6 cycle Morlet wavelets

Then, I extracted epoched data for each trial. The epochs were defined as follows:

* Start time: 3000 ms before teleporter entry
* End time: 3000 ms after teleporter exit

I calculated the power at each frequency at every time point of the epoch, using
mirrored buffers (length = 3 cycles at 1 Hz) to avoid edge artifacts, again
using 6 cycle Morlet wavelets.

This timeseries data was then amplitude-thresholded using the value at 95% of the
distribution calculated earlier. This returns a binary vector which indicates
whether power exceeded the amplitude threshold at each time point. This vector
was then duration-thresholded (n >= 3 cycles) so that only episodes lasting at
least 3 cycles are retained. The result is a binary vector showing times when
significant oscillatory episodes are present.

This vector was further processed to identify the onset and offset of each
episode. The onset was defined as the first time point of the episode; if an
episode was present at the first time point of the trial (i.e., the trial started
on an ongoing oscillation), then the onset was set as the first time point of the
trial. The offset was defined as the last time point of the episode; if an 
episode was present at the last time point of the trial (i.e., the trial ended
on an ongoing oscillation), then the offset was set as the last time point of the
trial.

Thus, this analysis output the following data for each electrode/frequency/trial:

* Binary vector indicating whether a significant episode was present at each 
time point of the trial
* Vector of onset times
* Vector of offset times

## Group Analysis of Episode Timing

To identify when oscillations were present across the group, I concatenated all
of the binary vectors of the same Time Type (NT/FT) row-wise and took the mean 
at each time point. Because these are binary vectors, the mean indicates the 
proportion of time episodes were present across all trials/electrodes.

The plots below show that pepisode is consistent across the entire trial in all
frequency bands. Edge effects can be seen at the beginning and end, with shorter
effects at higher frequencies. Vertical lines indicate times of teleporter entry
and exit.

```{r, echo = FALSE, message = FALSE, fig.width=20, fig.height=10}

library(dplyr)
library(ggplot2)
library(ggthemes)
library(grid)

setwd('~/Documents/MATLAB/iEEG/Group Analysis/Pepisode_Timing_Onset_Offset/')
load('Rda/allAnalyzedData.Rda')

# plot group data
timepointMarkers <- data.frame(x = c(0, 1830, 0, 2830), TrialTimeType = c("NT", "NT", "FT", "FT"))
lineData <- episodeData %>%
  group_by(FrequencyBand, TrialTimeType, Time) %>%
  summarise(GroupMean = mean(Mean), GroupSEM = sd(Mean) / sqrt(n())) 
linePlot <- lineData %>%
  ggplot(aes(x = Time, y = GroupMean, ymin = GroupMean - GroupSEM, ymax = GroupMean + GroupSEM)) + 
  geom_ribbon(color = "lightskyblue", fill = "lightskyblue") +
  geom_line(color = "steelblue4") +
  geom_vline(aes(xintercept = x, linetype = "dashed"), timepointMarkers) +
  facet_grid(FrequencyBand ~ TrialTimeType, scales = "free_x") +
  theme_few() +
  labs(y = "Mean Pepisode", title = "Pepisode Over Time") +
  theme(text = element_text(size = 24),
        axis.text = element_text(size = 18),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 1),
        panel.margin = unit(1, "lines"))
print(linePlot)

```

## Electrode-wise Analysis of Episode Timing

The above results, showing flat pepisode, could have occurred for two reasons:

* Pepisode is consistent over time within each electrode
* Pepisode peaks at a particular time within each electrode, but these peaks are
not consistent across electrodes

To adjudicate between these two possibilities, I plotted the pepisode timing
values separately for each electrode. Below I show just one of these plots, which
depicts theta pepisode for NT trials. All the other plots for the other frequency
bands and time types can be found in the Figures folder.

The individual electrode results appear to support the 1st conclusion: pepisode
is relatively consistent over time within an electrode. The individual electrode
data is noisier than the group data since it contains many fewer observations but
appears to be relatively flat for each electrode and frequency (with a few
exceptions -- see for example Delta FT).

```{r, echo = FALSE, message = FALSE, fig.width=30, fig.height=15}

# plot theta data for NT trials
lineNtData <- episodeData %>%
    filter(FrequencyBand == "Theta" & TrialTimeType == "NT") %>%
    ggplot(aes(x = Time, y = Mean, ymin = Mean - SEM, ymax = Mean + SEM)) + 
    geom_ribbon(color = "lightskyblue", fill = "lightskyblue") +
    geom_line(color = "steelblue4") +
    geom_vline(xintercept = 0) +
    geom_vline(xintercept = 1830) +
    facet_wrap(~ElectrodeID) +
    theme_few() +
    labs(y = "Mean Pepisode", title = "Theta Pepisode Over Time (NT Trials)") +
    theme(text = element_text(size = 24),
          axis.text = element_text(size = 18),
          axis.title.x = element_text(vjust = -0.5),
          axis.title.y = element_text(vjust = 1),
          panel.margin = unit(1, "lines"))
print(lineNtData)

```

## Group Analysis of Pepisode Onsets and Offsets

To identify when pepisodes tend to begin and end, I plotted the onset and offset
times in histogram form, as shown below. Each bin is 261 ms. Vertical lines indicate times of
teleporter entry and exit. The histograms show that pepisode onsets and offsets
are evenly distributed across the trial, with no apparent peaks at the times of
teleporter entry or exit. Peaks can be seen however in the first bin of the 
onsets (many trials begin on an oscillation) and the last bin of the offsets
(many trials end on an oscillation).

```{r, echo = FALSE, message = FALSE, fig.width=20, fig.height=10}

# plot group onset data
onsetData <- onOffData %>%
  ungroup() %>%
  filter(ObservationType == "Onset") %>%
  ggplot(aes(x = Time)) +
  geom_histogram(colour = "steelblue4", fill = "lightskyblue", binwidth = 261) +
  facet_grid(FrequencyBand ~ TrialTimeType, scales = "free") +
  geom_vline(aes(xintercept = x), timepointMarkers) +
  theme_few() +
  theme(text = element_text(size = 24),
        panel.margin = unit(1, "lines")) +
  labs(y = "# of Episodes", title = "Pepisode Onset Times")
print(onsetData)

# plot group offset data
offsetData <- onOffData %>%
  ungroup() %>%
  filter(ObservationType == "Offset") %>%
  ggplot(aes(x = Time)) +
  geom_histogram(colour = "steelblue4", fill = "lightskyblue", binwidth = 261) +
  facet_grid(FrequencyBand ~ TrialTimeType, scales = "free") +
  geom_vline(aes(xintercept = x), timepointMarkers) +
  theme_few() +
  theme(text = element_text(size = 24),
        panel.margin = unit(1, "lines")) +
  labs(y = "# of Episodes", title = "Pepisode Offset Times")
print(offsetData)

```

## Electrode-wise Analysis of Pepisode Onsets and Offsets

Much like the episode timing, the onsets and offsets could appear flat either
because each electrode's distribution is flat or because each electrode shows
unique peaks in the distribution. Below, I've plotted the onsets and offsets for
the theta band NT trials.

```{r, echo = FALSE, message = FALSE, fig.width=30, fig.height=15}



```