---
title: "Power Two-Way Repeated Measures ANOVA Results"
author: Lindsay Vass
date: 27 May 2015
output: html_document
---

# Calculation of Power

Six-cycle wavelets were used to estimate the power at each frequency and time
point of the cleaned data set. I then took the log(Power) and z-scored these
values within each electrode and frequency. All plotted values are of these
z-scored log(Power) values.

* 31 log-spaced frequencies between 1 Hz and 181 Hz
* 6 cycle wavelet for power estimation

I then extracted the mean power value for each trial in a series of 
contiguous time bins that surrounded the teleportation event. These bins were
defined as follows (t0 = time of teleporter entry):

* -3000 : -2001 ms
* -2000 : -1001 ms
* -1000 : 0 ms
* Time in Teleporter (0 : 1830 ms for short time; 0 : 2830 ms for long time)
* Exit time : Exit time + 1000 ms
* Exit time + 1001 ms : Exit time + 2000 ms
* Exit time + 2001 ms : Exit time + 3000 ms

# Repeated Measures ANOVA and permutation testing

Once the power values were extracted, I submitted them to two sets of 
two-way repeated measures ANOVAs:

* Space Condition x Timepoint
* Time Condition x Timepoint

After calculating the F statistics of the observed data, I built a distribution
of F statistics for each ANOVA in order to calculate the true P value. The 
current analysis used 10000 iterations to build each distribution. In each 
section below, I plot the data from a given frequency band, and display the 
corresponding F statistics and the estimated P values.

``` {r, echo = FALSE, message = FALSE}
library(dplyr)
library(ggplot2)

setwd('/Users/Lindsay/Documents/MATLAB/iEEG/Group Analysis/Power/')
load("All_Subjects_Power_CLEAN_3sBuffer2015-05-27.Rda")
load("All_Subjects_Power_CLEAN_3sBuffer_Two_Way_RMANOVA_2015-05-27.Rda")

# Set up data by TIMEPOINT ---------------------------------------------------
meanByTimePoint <- allPowerData %>%
  filter(is.na(Power) == FALSE) %>% # UCDMC15 TeleporterB is NaN for NSFT
  select(TimePoint,Power:FrequencyBand) %>%
  group_by(ElectrodeID, FrequencyBand,TimePoint) 

# We want our error bars to ultimately reflect the within-subject variability rather than between-subject
# To do this, we'll use the methods in Cousineau (2005) Tutorials in Quantitative Methods for Psychology and
# Morey (2008) [Same Journal], which corrects the bias in the Cousineau method
# Essentially, we will calculate normalized values for each observation, in which we take the original observation,
# subtract the electrode mean, add the group mean, and correct for the # of conditions

# calculate the mean across conditions for each electrode for each frequency
elecMean <- meanByTimePoint %>%
  group_by(ElectrodeID, FrequencyBand) %>%
  summarise(EMean = mean(Power))

# calculate the mean across conditions across electrodes for each frequency
groupMean <- meanByTimePoint %>%
  group_by(FrequencyBand) %>%
  summarise(GMean = mean(Power))

# now we'll normalize the observations by taking into account the electrode and group means
# to do this, we'll need to join our tables together
normByTimePoint <- inner_join(meanByTimePoint, elecMean)
normByTimePoint <- inner_join(normByTimePoint, groupMean) %>%
  mutate(NormPower = Power - EMean + GMean) %>%
  filter(is.na(Power) == FALSE) %>%
  group_by(FrequencyBand, TimePoint) %>%
  summarise(NormVar = var(NormPower)) %>%
  mutate(NormVarUnbias = NormVar * (nlevels(allPowerData$TimePoint)/(nlevels(allPowerData$TimePoint) - 1)),
         NormSEM = (sqrt(NormVarUnbias) / sqrt(nlevels(allPowerData$ElectrodeID)))) %>%
  select(FrequencyBand, TimePoint, NormSEM)

# now summarise the original data and join with the normalized SEM
meanByTimePoint <- meanByTimePoint %>%
  group_by(FrequencyBand, TimePoint) %>%
  summarise(Power = mean(Power)) %>%
  inner_join(normByTimePoint)

# Summarize the data by SPACE condition ---------------------------------------

meanBySpace <- allPowerData %>%
  filter(is.na(Power) == FALSE) %>% # UCDMC15 TeleporterB is NaN for EDF1 NSFT
  select(TrialSpaceType,TimePoint,Power:FrequencyBand) %>%
  group_by(ElectrodeID, FrequencyBand,TrialSpaceType) 

# We want our error bars to ultimately reflect the within-subject variability rather than between-subject
# To do this, we'll use the methods in Cousineau (2005) Tutorials in Quantitative Methods for Psychology and
# Morey (2008) [Same Journal], which corrects the bias in the Cousineau method
# Essentially, we will calculate normalized values for each observation, in which we take the original observation,
# subtract the electrode mean, add the group mean, and correct for the # of conditions

# calculate the mean across conditions for each electrode for each frequency
elecMean <- meanBySpace %>%
  group_by(ElectrodeID, FrequencyBand, TimePoint, TrialSpaceType) %>%
  summarise(EMean = mean(Power))

# calculate the mean across conditions across electrodes for each frequency
groupMean <- meanBySpace %>%
  group_by(FrequencyBand, TimePoint, TrialSpaceType) %>%
  summarise(GMean = mean(Power))

# now we'll normalize the observations by taking into account the electrode and group means
# to do this, we'll need to join our tables together
normBySpace <- inner_join(meanBySpace, elecMean)
normBySpace <- inner_join(normBySpace, groupMean) %>%
  mutate(NormPower = Power - EMean + GMean) %>%
  group_by(FrequencyBand, TimePoint, TrialSpaceType) %>%
  summarise(NormVar = var(NormPower)) %>%
  mutate(NormVarUnbias = NormVar * (nlevels(allPowerData$TrialSpaceType))/(nlevels(allPowerData$TrialSpaceType) - 1),
         NormSEM = (sqrt(NormVarUnbias) / sqrt(nlevels(allPowerData$ElectrodeID)))) %>%
  select(FrequencyBand, TimePoint, TrialSpaceType, NormSEM)

# now summarise the original data and join with the normalized SEM
meanBySpace <- meanBySpace %>%
  group_by(FrequencyBand, TrialSpaceType, TimePoint) %>%
  summarise(Power = mean(Power)) %>%
  inner_join(normByTimePoint)

# Summarize the data by TIME condition ----------------------------------------
meanByTime <- allPowerData %>%
  filter(is.na(Power) == FALSE) %>% # UCDMC15 TeleporterB is NaN for EDF1 NSFT
  select(TrialTimeType,TimePoint,Power:FrequencyBand) %>%
  group_by(ElectrodeID, FrequencyBand,TrialTimeType) 

# We want our error bars to ultimately reflect the within-subject variability rather than between-subject
# To do this, we'll use the methods in Cousineau (2005) Tutorials in Quantitative Methods for Psychology and
# Morey (2008) [Same Journal], which corrects the bias in the Cousineau method
# Essentially, we will calculate normalized values for each observation, in which we take the original observation,
# subtract the electrode mean, add the group mean, and correct for the # of conditions

# calculate the mean across conditions for each electrode for each frequency
elecMean <- meanByTime %>%
  group_by(ElectrodeID, FrequencyBand, TimePoint, TrialTimeType) %>%
  summarise(EMean = mean(Power))

# calculate the mean across conditions across electrodes for each frequency
groupMean <- meanByTime %>%
  group_by(FrequencyBand, TimePoint, TrialTimeType) %>%
  summarise(GMean = mean(Power))

# now we'll normalize the observations by taking into account the electrode and group means
# to do this, we'll need to join our tables together
normByTime <- inner_join(meanByTime, elecMean)
normByTime <- inner_join(normByTime, groupMean) %>%
  mutate(NormPower = Power - EMean + GMean) %>%
  group_by(FrequencyBand, TimePoint, TrialTimeType) %>%
  summarise(NormVar = var(NormPower)) %>%
  mutate(NormVarUnbias = NormVar * (nlevels(allPowerData$TrialTimeType))/(nlevels(allPowerData$TrialTimeType) - 1),
         NormSEM = (sqrt(NormVarUnbias) / sqrt(nlevels(allPowerData$ElectrodeID)))) %>%
  select(FrequencyBand, TimePoint, TrialTimeType, NormSEM)

# now summarise the original data and join with the normalized SEM
meanByTime <- meanByTime %>%
  group_by(FrequencyBand, TrialTimeType, TimePoint) %>%
  summarise(Power = mean(Power)) %>%
  inner_join(normByTimePoint)
```

```{r, echo = FALSE, comment = ""}

for (thisFreq in 1:nlevels(allPowerData$FrequencyBand)) {
  
  thisFreqName <- levels(allPowerData$FrequencyBand)[thisFreq]
  
  # Plot the data
  meanBySpace <- meanBySpace %>%
    filter(FrequencyBand == thisFreqName) 
  plot1 <-  ggplot(data = meanBySpace,
                   aes(x = TimePoint,
                       y = Power,
                       ymin = Power - NormSEM,
                       ymax = Power + NormSEM,
                       color = TrialSpaceType)) +
    geom_point(size = 4) +
    geom_pointrange() +
    ggtitle("Power: Space x Timepoint")
  print(plot1)

  
  # Output the stats
  results1 <- anovaResults %>% 
    filter(FreqBand == thisFreqName & ANOVA == "SpaceXTimepoint") %>% 
    select(Contrast:Pcorr)
  print(results1)

  
  # Plot the data
  meanByTime <- meanByTime %>%
    filter(FrequencyBand == thisFreqName)
  plot2 <-  ggplot(data = meanByTime,
                   aes(x = TimePoint,
                       y = Power,
                       ymin = Power - NormSEM,
                       ymax = Power + NormSEM,
                       color = TrialTimeType)) +
    geom_point(size = 4) +
    geom_pointrange() +
    ggtitle("Power: Time x Timepoint")
  print(plot2)
  
  # Output the stats
  results2 <- anovaResults %>% 
    filter(FreqBand == thisFreqName & ANOVA == "TimeXTimepoint") %>% 
    select(Contrast:Pcorr)
  print(results2)
  
  }


```