---
title: "Pepisode Two-Way Repeated Measures ANOVA Results"
author: Lindsay Vass
date: 26 May 2015
output: html_document
---

# Calculation of Pepisode

The cleaned data was submitted to a pepisode analysis which consisted of two 
parts. In the first part, power at each frequency was estimated at each time
point of the entire navigation period when subjects were searching for stores.
This resulted in a distribution of power values at each frequency across the
experiment. In the second part of the analysis, the script determined for each
time point whether there was an ongoing oscillation present, defined as power
that exceeded both an amplitude and duration threshold, as indicated below.

* 31 log-spaced frequencies between 1 Hz and 181 Hz
* 6 cycle wavelet for power estimation
* Amplitude threshold = 95% of the distribution
* Duration threshold = 3 cycles

I then extracted the mean pepisode value for each trial in a series of 
contiguous time bins that surrounded the teleportation event. These bins were
defined as follows (t0 = time of teleporter entry):

* -3000 : -2001 ms
* -2000 : -1001 ms
* -1000 : 0 ms
* Time in Teleporter (0 : 1830 ms for short time; 0 : 2830 ms for long time)
* Exit time : Exit time + 1000 ms
* Exit time + 1001 ms : Exit time + 2000 ms
* Exit time + 2001 ms : Exit time + 3000 ms

# Repeated Measures ANOVA and permutation testing

Once the pepisode values were extracted, I submitted them to two sets of 
two-way repeated measures ANOVAs:

* Space Condition x Timepoint
* Time Condition x Timepoint

After calculating the F statistics of the observed data, I built a distribution
of F statistics for each ANOVA in order to calculate the true P value since 
pepisode values are not normally distributed. The current analysis used 10000
iterations to build each distribution. In each section below, I plot the data
from a given frequency band, and display the corresponding F statistics and the
estimated P values.

``` {r, echo = FALSE, message = FALSE}
library(dplyr)
library(ggplot2)

setwd('/Users/Lindsay/Documents/MATLAB/iEEG/Group Analysis/Pepisode/')
load("All_Subjects_Pepisode_CLEAN_3sBuffer2015-05-22.Rda")
load("All_Subjects_Pepisode_CLEAN_3sBuffer_Two_Way_RMANOVA_2015-05-26.Rda")

# Set up data by TIMEPOINT ---------------------------------------------------
meanByTimePoint <- allPepisodeData %>%
  filter(is.na(Pepisode) == FALSE) %>% # UCDMC15 TeleporterB is NaN for NSFT
  select(TimePoint,Pepisode:FrequencyBand) %>%
  group_by(ElectrodeID, FrequencyBand,TimePoint) 

# We want our error bars to ultimately reflect the within-subject variability rather than between-subject
# To do this, we'll use the methods in Cousineau (2005) Tutorials in Quantitative Methods for Psychology and
# Morey (2008) [Same Journal], which corrects the bias in the Cousineau method
# Essentially, we will calculate normalized values for each observation, in which we take the original observation,
# subtract the electrode mean, add the group mean, and correct for the # of conditions

# calculate the mean across conditions for each electrode for each frequency
elecMean <- meanByTimePoint %>%
  group_by(ElectrodeID, FrequencyBand) %>%
  summarise(EMean = mean(Pepisode))

# calculate the mean across conditions across electrodes for each frequency
groupMean <- meanByTimePoint %>%
  group_by(FrequencyBand) %>%
  summarise(GMean = mean(Pepisode))

# now we'll normalize the observations by taking into account the electrode and group means
# to do this, we'll need to join our tables together
normByTimePoint <- inner_join(meanByTimePoint, elecMean)
normByTimePoint <- inner_join(normByTimePoint, groupMean) %>%
  mutate(NormPepisode = Pepisode - EMean + GMean) %>%
  filter(is.na(Pepisode) == FALSE) %>%
  group_by(FrequencyBand, TimePoint) %>%
  summarise(NormVar = var(NormPepisode)) %>%
  mutate(NormVarUnbias = NormVar * (nlevels(allPepisodeData$TimePoint)/(nlevels(allPepisodeData$TimePoint) - 1)),
         NormSEM = (sqrt(NormVarUnbias) / sqrt(nlevels(allPepisodeData$ElectrodeID)))) %>%
  select(FrequencyBand, TimePoint, NormSEM)

# now summarise the original data and join with the normalized SEM
meanByTimePoint <- meanByTimePoint %>%
  group_by(FrequencyBand, TimePoint) %>%
  summarise(Pepisode = mean(Pepisode)) %>%
  inner_join(normByTimePoint)

# Summarize the data by SPACE condition ---------------------------------------

meanBySpace <- allPepisodeData %>%
  filter(is.na(Pepisode) == FALSE) %>% # UCDMC15 TeleporterB is NaN for EDF1 NSFT
  select(TrialSpaceType,TimePoint,Pepisode:FrequencyBand) %>%
  group_by(ElectrodeID, FrequencyBand,TrialSpaceType) 

# We want our error bars to ultimately reflect the within-subject variability rather than between-subject
# To do this, we'll use the methods in Cousineau (2005) Tutorials in Quantitative Methods for Psychology and
# Morey (2008) [Same Journal], which corrects the bias in the Cousineau method
# Essentially, we will calculate normalized values for each observation, in which we take the original observation,
# subtract the electrode mean, add the group mean, and correct for the # of conditions

# calculate the mean across conditions for each electrode for each frequency
elecMean <- meanBySpace %>%
  group_by(ElectrodeID, FrequencyBand, TimePoint, TrialSpaceType) %>%
  summarise(EMean = mean(Pepisode))

# calculate the mean across conditions across electrodes for each frequency
groupMean <- meanBySpace %>%
  group_by(FrequencyBand, TimePoint, TrialSpaceType) %>%
  summarise(GMean = mean(Pepisode))

# now we'll normalize the observations by taking into account the electrode and group means
# to do this, we'll need to join our tables together
normBySpace <- inner_join(meanBySpace, elecMean)
normBySpace <- inner_join(normBySpace, groupMean) %>%
  mutate(NormPepisode = Pepisode - EMean + GMean) %>%
  group_by(FrequencyBand, TimePoint, TrialSpaceType) %>%
  summarise(NormVar = var(NormPepisode)) %>%
  mutate(NormVarUnbias = NormVar * (nlevels(allPepisodeData$TrialSpaceType))/(nlevels(allPepisodeData$TrialSpaceType) - 1),
         NormSEM = (sqrt(NormVarUnbias) / sqrt(nlevels(allPepisodeData$ElectrodeID)))) %>%
  select(FrequencyBand, TimePoint, TrialSpaceType, NormSEM)

# now summarise the original data and join with the normalized SEM
meanBySpace <- meanBySpace %>%
  group_by(FrequencyBand, TrialSpaceType, TimePoint) %>%
  summarise(Pepisode = mean(Pepisode)) %>%
  inner_join(normByTimePoint)

# Summarize the data by TIME condition ----------------------------------------
meanByTime <- allPepisodeData %>%
  filter(is.na(Pepisode) == FALSE) %>% # UCDMC15 TeleporterB is NaN for EDF1 NSFT
  select(TrialTimeType,TimePoint,Pepisode:FrequencyBand) %>%
  group_by(ElectrodeID, FrequencyBand,TrialTimeType) 

# We want our error bars to ultimately reflect the within-subject variability rather than between-subject
# To do this, we'll use the methods in Cousineau (2005) Tutorials in Quantitative Methods for Psychology and
# Morey (2008) [Same Journal], which corrects the bias in the Cousineau method
# Essentially, we will calculate normalized values for each observation, in which we take the original observation,
# subtract the electrode mean, add the group mean, and correct for the # of conditions

# calculate the mean across conditions for each electrode for each frequency
elecMean <- meanByTime %>%
  group_by(ElectrodeID, FrequencyBand, TimePoint, TrialTimeType) %>%
  summarise(EMean = mean(Pepisode))

# calculate the mean across conditions across electrodes for each frequency
groupMean <- meanByTime %>%
  group_by(FrequencyBand, TimePoint, TrialTimeType) %>%
  summarise(GMean = mean(Pepisode))

# now we'll normalize the observations by taking into account the electrode and group means
# to do this, we'll need to join our tables together
normByTime <- inner_join(meanByTime, elecMean)
normByTime <- inner_join(normByTime, groupMean) %>%
  mutate(NormPepisode = Pepisode - EMean + GMean) %>%
  group_by(FrequencyBand, TimePoint, TrialTimeType) %>%
  summarise(NormVar = var(NormPepisode)) %>%
  mutate(NormVarUnbias = NormVar * (nlevels(allPepisodeData$TrialTimeType))/(nlevels(allPepisodeData$TrialTimeType) - 1),
         NormSEM = (sqrt(NormVarUnbias) / sqrt(nlevels(allPepisodeData$ElectrodeID)))) %>%
  select(FrequencyBand, TimePoint, TrialTimeType, NormSEM)

# now summarise the original data and join with the normalized SEM
meanByTime <- meanByTime %>%
  group_by(FrequencyBand, TrialTimeType, TimePoint) %>%
  summarise(Pepisode = mean(Pepisode)) %>%
  inner_join(normByTimePoint)
```

# Delta

## Space x Timepoint

```{r, echo = FALSE}

# Plot the data
meanBySpace %>%
  filter(FrequencyBand == "Delta") %>%
  ggplot(aes(x = TimePoint,
             y = Pepisode,
             ymin = Pepisode - NormSEM,
             ymax = Pepisode + NormSEM,
             color = TrialSpaceType)) +
  geom_point(size = 4) +
  geom_pointrange() +
  ggtitle("Pepisode: Space x Timepoint")

# Output the stats
anovaResults %>% 
  filter(FreqBand == "Delta" & ANOVA == "SpaceXTimepoint") %>% 
  select(Contrast:Pcorr)

```

## Time x Timepoint

```{r, echo = FALSE}

# Plot the data
meanByTime %>%
  filter(FrequencyBand == "Delta") %>%
  ggplot(aes(x = TimePoint,
             y = Pepisode,
             ymin = Pepisode - NormSEM,
             ymax = Pepisode + NormSEM,
             color = TrialTimeType)) +
  geom_point(size = 4) +
  geom_pointrange() +
  ggtitle("Pepisode: Time x Timepoint")

# Output the stats
anovaResults %>% 
  filter(FreqBand == "Delta" & ANOVA == "TimeXTimepoint") %>% 
  select(Contrast:Pcorr)

```

# Theta

## Space x Timepoint

```{r, echo = FALSE}

# Plot the data
meanBySpace %>%
  filter(FrequencyBand == "Theta") %>%
  ggplot(aes(x = TimePoint,
             y = Pepisode,
             ymin = Pepisode - NormSEM,
             ymax = Pepisode + NormSEM,
             color = TrialSpaceType)) +
  geom_point(size = 4) +
  geom_pointrange() +
  ggtitle("Pepisode: Space x Timepoint")

# Output the stats
anovaResults %>% 
  filter(FreqBand == "Theta" & ANOVA == "SpaceXTimepoint") %>% 
  select(Contrast:Pcorr)

```

## Time x Timepoint

```{r, echo = FALSE}

# Plot the data
meanByTime %>%
  filter(FrequencyBand == "Theta") %>%
  ggplot(aes(x = TimePoint,
             y = Pepisode,
             ymin = Pepisode - NormSEM,
             ymax = Pepisode + NormSEM,
             color = TrialTimeType)) +
  geom_point(size = 4) +
  geom_pointrange() +
  ggtitle("Pepisode: Time x Timepoint")

# Output the stats
anovaResults %>% 
  filter(FreqBand == "Theta" & ANOVA == "TimeXTimepoint") %>% 
  select(Contrast:Pcorr)

```

# Alpha

## Space x Timepoint

```{r, echo = FALSE}

# Plot the data
meanBySpace %>%
  filter(FrequencyBand == "Alpha") %>%
  ggplot(aes(x = TimePoint,
             y = Pepisode,
             ymin = Pepisode - NormSEM,
             ymax = Pepisode + NormSEM,
             color = TrialSpaceType)) +
  geom_point(size = 4) +
  geom_pointrange() +
  ggtitle("Pepisode: Space x Timepoint")

# Output the stats
anovaResults %>% 
  filter(FreqBand == "Alpha" & ANOVA == "SpaceXTimepoint") %>% 
  select(Contrast:Pcorr)

```

## Time x Timepoint

```{r, echo = FALSE}

# Plot the data
meanByTime %>%
  filter(FrequencyBand == "Alpha") %>%
  ggplot(aes(x = TimePoint,
             y = Pepisode,
             ymin = Pepisode - NormSEM,
             ymax = Pepisode + NormSEM,
             color = TrialTimeType)) +
  geom_point(size = 4) +
  geom_pointrange() +
  ggtitle("Pepisode: Time x Timepoint")

# Output the stats
anovaResults %>% 
  filter(FreqBand == "Alpha" & ANOVA == "TimeXTimepoint") %>% 
  select(Contrast:Pcorr)

```

# Beta

## Space x Timepoint

```{r, echo = FALSE}

# Plot the data
meanBySpace %>%
  filter(FrequencyBand == "Beta") %>%
  ggplot(aes(x = TimePoint,
             y = Pepisode,
             ymin = Pepisode - NormSEM,
             ymax = Pepisode + NormSEM,
             color = TrialSpaceType)) +
  geom_point(size = 4) +
  geom_pointrange() +
  ggtitle("Pepisode: Space x Timepoint")

# Output the stats
anovaResults %>% 
  filter(FreqBand == "Beta" & ANOVA == "SpaceXTimepoint") %>% 
  select(Contrast:Pcorr)

```

## Time x Timepoint

```{r, echo = FALSE}

# Plot the data
meanByTime %>%
  filter(FrequencyBand == "Beta") %>%
  ggplot(aes(x = TimePoint,
             y = Pepisode,
             ymin = Pepisode - NormSEM,
             ymax = Pepisode + NormSEM,
             color = TrialTimeType)) +
  geom_point(size = 4) +
  geom_pointrange() +
  ggtitle("Pepisode: Time x Timepoint")

# Output the stats
anovaResults %>% 
  filter(FreqBand == "Beta" & ANOVA == "TimeXTimepoint") %>% 
  select(Contrast:Pcorr)

```

# Gamma

## Space x Timepoint

```{r, echo = FALSE}

# Plot the data
meanBySpace %>%
  filter(FrequencyBand == "Gamma") %>%
  ggplot(aes(x = TimePoint,
             y = Pepisode,
             ymin = Pepisode - NormSEM,
             ymax = Pepisode + NormSEM,
             color = TrialSpaceType)) +
  geom_point(size = 4) +
  geom_pointrange() +
  ggtitle("Pepisode: Space x Timepoint")

# Output the stats
anovaResults %>% 
  filter(FreqBand == "Gamma" & ANOVA == "SpaceXTimepoint") %>% 
  select(Contrast:Pcorr)

```

## Time x Timepoint

```{r, echo = FALSE}

# Plot the data
meanByTime %>%
  filter(FrequencyBand == "Gamma") %>%
  ggplot(aes(x = TimePoint,
             y = Pepisode,
             ymin = Pepisode - NormSEM,
             ymax = Pepisode + NormSEM,
             color = TrialTimeType)) +
  geom_point(size = 4) +
  geom_pointrange() +
  ggtitle("Pepisode: Time x Timepoint")

# Output the stats
anovaResults %>% 
  filter(FreqBand == "Gamma" & ANOVA == "TimeXTimepoint") %>% 
  select(Contrast:Pcorr)

```



